Local to global N input modelling

### 2025-07-23 manuscript.qmd
Docs fixup and render

This document was last updated at `r format(Sys.time(), "%H:%M")` on `r format(Sys.time(), "%d %b, %Y")`.

@@

worldmap <- ne_countries(scale = "medium", returnclass = "sf") %>% st_transform(crs = "+proj=robin")
graticules <- st_graticule(worldmap, lon = seq(-180, 180, 30), lat = seq(-90, 90, 30))

inset_boxes <- list(   # lonmin, lonmax, latmin, latmax
  CAN1 = c(-132, -122, 47.75, 54.25),
  CAN2 = c(-70, -54, 43, 48.5),
  EUR = c(-26, 30, 51, 72),
  CHI = c(-77.5, -62.5, -56, -25),
  AUS = c(144, 149.5, -44, -39.75)
)
labels_spec <- c(CAN1 = "bottom_left_outside", CAN2 = "bottom_right_outside", EUR = "bottom_left_outside", CHI = "top_left_outside", AUS = "bottom_left_outside")
labels_offset <- c(CAN1 = 1, CAN2 = 2, EUR = 1.75, CHI = 1.75, AUS = 3.5)

box_data <- create_boxes(box_list = inset_boxes, # see function create_boxes above
                         label_positions = unname(labels_spec),
                         offset_deg = unname(labels_offset))
boxes_robinson <- st_transform(box_data$boxes, crs = "+proj=robin")
labels_robinson <- st_transform(box_data$labels, crs = "+proj=robin")
farm_temp_robinson <- st_sf(farm_temp_means) %>% st_transform(crs = "+proj=robin")

p_big_map <- ggplot() +
  geom_sf(data = graticules, color = "gray80", size = 0.3) +
  geom_sf(data = worldmap, fill = "white", color = "dimgray") +
  geom_sf(data = farm_temp_robinson, aes(color = mean_temp), size = 1) +
  coord_sf() +
  scale_color_viridis_c(option = "turbo") +
  guides(col = guide_colourbar(
    title = "Mean temperature (°C)",
    direction = "horizontal", position = "bottom",
    label.position = "bottom", title.position = "top", 
    title.vjust = 1, title.hjust = 0.5,
    frame.colour = "black", ticks.colour = "black",ticks.linewidth = 2,
    barwidth = 25,barheight = 1.5
  )) +
  geom_sf(data = boxes_robinson, fill = NA, color = "darkred", size = 1.65) +
  geom_sf_text(data = labels_robinson, aes(label = letter), 
               color = "darkred", size = 5, fontface = "bold", 
               hjust = 0.5, vjust = 0.5) +
  theme_void()

@@

no_margins <- function() {
  theme(legend.position = "none", 
        plot.margin = margin(0, 0, 0, 0),
        axis.title = element_blank())
}

inset_boxes_sm <- list(   # lonmin, lonmax, latmin, latmax
  CAN1 = c(-132, -122, 48.5, 54),
  CAN2 = c(-70, -54, 43, 48.5),
  EUR = c(-23, 28.5, 52, 71),
  CHI = c(-78, -62, -55.5, -27),
  AUS = c(144, 149, -43.75, -40.75)
)

inset_boxes_sm <- map(inset_boxes_sm, function(bx) {
  list(
    xlims = bx[1:2],
    ylims = bx[3:4],
    labx = bx[1],
    laby = bx[4]
    )
})

worldmap <- ne_countries(scale = "large", returnclass = "sf")
p1 <- ggplot() +
  geom_sf(data = worldmap, fill = "white", color = "dimgray") +
  geom_sf(data = farm_temp_means, aes(color = mean_temp, geometry = geometry), size = 2) +
  coord_sf() +
  scale_color_viridis_c(option = "turbo") +
  guides(col = guide_colourbar(
    title = "Mean temperature (°C)",
    direction = "vertical", position = "right",
    label.position = "right", title.position = "top", 
    title.vjust = 1, title.hjust = 0.5,
    frame.colour = "black", ticks.colour = "black",ticks.linewidth = 2,
    barwidth = 1.5, barheight = 20
  )) +
  prettyplot() +
  labs(y = "Latitude", x = "Longitude")

@@

## Model parameterisation

<!-- Salmon were grown for 548 days and were initialised in Spring. This was on day `r unique(farm_coords$t_start[farm_coords$lat < 0])` for farms in the southern hemisphere and day `r unique(farm_coords$t_start[farm_coords$lat > 0])` for farms in the northern hemisphere. Cohorts were initialised annually, such that there was an overlap of 183 days where one cohort was nearing harvest and another was being initialised. -->

```{r salmon params}
#| code-summary: Get parameter and sensitivity data
#| eval: true

params <- file.path(output_species_data_path, "sens_params.qs") %>% qs::qread()

param_names <- tibble::tribble(
  ~name,        ~lab,
  "alpha",      bquote(alpha),
  "epsprot",    bquote(epsilon["P"]),
  "epslip",     bquote(epsilon["L"]),
  "epscarb",    bquote(epsilon["C"]),
  "epsO2",      bquote(epsilon["O"[2]]),
  "pk",         bquote("pk"),
  "k0",         bquote("k"[0]),
  "m",          bquote("m"),
  "n",          bquote("n"),
  "betac",      bquote(beta*"C"),
  "Tma",        bquote("T"["max"]),
  "Toa",        bquote("T"["opt"]),
  "Taa",        bquote("T"["min"]),
  "omega",      bquote(omega),
  "a",          bquote("a"),
  "k",          bquote("k"),
  "eff",        bquote("eff"),
  "meanW",      bquote(bar("W")),
  "deltaW",     bquote(Delta*"W"), 
  "meanImax",   bquote(bar("I"["max"])),
  "deltaImax",  bquote(Delta*"I"["max"]),
  "overFmean",  bquote(bar("F")),
  "overFdelta", bquote(Delta*"F"),
  "mortmyt",    bquote("mort")
)
```

| Parameter | Value | Reference |
|:------:|:------:|:-----------:|
| $\alpha$             | `r unname(params['alpha'])`      | @broekhuizen_modelling_1994 |
| $\epsilon_{P}$       | `r unname(params['epsprot'])`    | @baldan_r_2018 |
| $\epsilon_{L}$       | `r unname(params['epslip'])`     | @baldan_r_2018 |
| $\epsilon_{C}$       | `r unname(params['epscarb'])`    | @baldan_r_2018 |
| $\epsilon_{O_2}$     | `r unname(params['epsO2'])`      | @elliott_energy_1975 |
| $p_k$                | `r unname(params['pk'])`         | @macnaughton_using_2019 |
| $k_0$                | `r unname(params['k0'])`         | @macnaughton_using_2019 |
| $m$                  | `r unname(params['m'])`          | @broekhuizen_modelling_1994 | 
| $n$                  | `r unname(params['n'])`          | @baldan_r_2018 |
| $\beta C$            | `r unname(params['betac'])`      | @jonsson_thermal_2001 |
| $T_{max}$            | `r unname(params['Tma'])`        | @poletto_unusual_2017 | 
| $T_{opt}$            | `r unname(params['Toa'])`        | @jonsson_thermal_2001 | 
| $T_{min}$            | `r unname(params['Taa'])`        | @vadboncoeur_lowering_2023 | 
| $\omega$             | `r unname(params['omega'])`      | @baldan_r_2018 |
| $a$                  | `r unname(params['a'])`          | @van_tien_development_2016 |
| $k$                  | `r unname(params['k'])`          | @van_tien_development_2016 | 
| $eff$                | `r unname(params['eff'])`        | @uglem_does_2020 | 
| $\overline{W}$       | `r unname(params['meanW'])`      | @dempsey_estimating_2023 | 
| $\Delta{W}$          | `r unname(params['deltaW'])`     | @dempsey_estimating_2023 | |
| $\overline{I_{max}}$ | `r unname(params['meanImax'])`   | @lall_nutrition_2009 | 
| $\Delta{I_{max}}$    | `r unname(params['deltaImax'])`  | @lall_nutrition_2009 | 
| $\overline{F}$       | `r unname(params['overFmean'])`  | |
| $\Delta{F}$          | `r unname(params['overFdelta'])` | |
| $mort$               | `r unname(params['mortmyt'])`    | @tvete_towards_2023 |
: Model growth parameters for Atlantic salmon (for double-checking numbers and reference list only) {#tbl-params}

### Parameter sensitivity

Parameter sensitivity was tested by varying each parameter in @tbl-params by $\pm 10$% and calculating sensitivity as $S_p = \frac{(P_{1.1} - P_{0.9})}{0.2 \times P_1}$, where $S$ is the sensitivity of the measured variable to changes in parameter $p$, and $P_{1.1}$, $P_{0.9}$, and $P_{1}$ are the value of the measured variable when the model is run with parameter $p$ at 110%, 90%, and 100% of its value respectively. 
All parameters in Table @tbl-params were tested. The population-level parameters ($\overline{W}$, $\Delta{W}$, $\overline{I_{max}}$, $\Delta{I_{max}}$, $\overline{F}$, $\Delta{F}$) were tested with a population of of 5000 on a random 10% of farms, while all other parameters were tested across all farms with a population of 10.

@@

## Model population and biomass

```{r raw biomass and weight data}
farmrun_files <- file.path(output_cohorts_data_path) %>% 
  list.files(full.names = T)

weight <- find_read(output_cohorts_data_path, "allfarms_comparisons_weight_stat")
biomass <- find_read(output_cohorts_data_path, "allfarms_comparisons_biomass_stat")
```

```{r fig-cohorts}
#| fig-cap: | 
#|   Example of how the overlapping cohorts combine into a single farm production curve (in this case biomass). The left plot shows the farm biomass produced by each cohort from initialisation to harvest, while the right plot shows the combined farm biomass over the course of a single production period of 548 days.
#| eval: true

fnms <- output_growth_data_path %>% 
  list.files(full.names = T) %>% 
  str_subset("reference")

df <- qs::qread(fnms[1])[["biomass_stat"]] %>% 
  as.data.frame() %>% 
  mutate(mean = set_units(mean, "g") %>% set_units("t"),
         sd = set_units(sd, "g") %>% set_units("t")) %>% 
  select(-farm_ID)

df <- rbind(
  df %>% mutate(cohort = 1),
  df %>% mutate(t = t+365, cohort = 2),
  df %>% mutate(t = t+730, cohort = 3)
) %>% 
  mutate(t = (t-min(t))+1,
         split = "Split by cohort")

min_t <- min(df$t[df$cohort == 2])
max_t <- max(df$t[df$cohort == 2])

df_2 <- df %>% 
  filter(t %in% min_t:max_t) %>% 
  mutate(sd = sd/mean) %>% 
  group_by(t) %>% 
  reframe(sd = sqrt(sum(sd^2)),
          mean = sum(mean),
          sd = sd*mean) %>% 
  mutate(cohort = "all",
         split = "Totalled")

df <- rbind(df, df_2) %>% 
  mutate(cohort = as.factor(cohort))

ggplot(df, aes(x = t, y = mean, ymin = mean-sd, ymax = mean+sd, colour = as.factor(cohort))) +
  geom_line(linewidth = 0.75) +
  geom_ribbon(linetype = "dotted", alpha = 0) +
  facet_grid(cols = vars(split)) +
  scale_x_continuous(breaks = seq(0, 1600, 120)) +
  scale_colour_manual(values = cohort_pal) +
  prettyplot() +
  labs(y = "Farm biomass", x = "Day of production")
```

```{r pop-biom-data}
coho_biom <- file.path(data_analysis_path, "biomass_produced_comparison.qs") %>% 
  qs::qread()
harv_size <- file.path(output_farm_data_path, "farm_harvest_size.qs") %>% 
  qread()

coho_biom_summ <- coho_biom %>% 
  mutate(perc_diff = 100*(mean - tonnes_per_farm)/tonnes_per_farm)
```

Each farm's final population was calculated by dividing the production tonnage required by the average modelled individual harvest size for each farm (when feed reference feed). A population timeseries was the generated using the standard mortality rate (`r unname(params['mortmyt'])` d$^{-1}$). Using this method, a farm's 'population' is not a good representation of the actual number of individuals being farmed, but it ensures each farm location is producing the correct tonnage while allowing comparisons between the different feeds.

```{r fig-harvest-size, class.chunk = "aside"}
#| fig-cap: |
#|   Frequency distribution of harvest sizes used to create the population timeseries for each farm, as predicted by the fish-growth model.

harv_size %>% 
  mutate(weight = weight %>% set_units("g") %>% set_units("kg")) %>% 
  select(c(farm_ID, weight)) %>% 
  ggplot(aes(x = weight)) +
  geom_histogram(binwidth = 0.25, colour = "black", fill = "salmon", alpha = 0.75) +
  scale_y_continuous(limits = c(0,500)) +
  labs(y = "Frequency", x = "Harvest weight") +
  prettyplot() +
  theme(aspect.ratio = 0.75)
```

By this method, farm biomass as predicted by the model slightly overestimated farm tonnage. Predicted farm biomass was `r round(min(coho_biom_summ$perc_diff),1)`-`r round(max(coho_biom_summ$perc_diff),1)` higher than tonnage produced (@fig-biomass-produced). This variation was not correlated with mean farm temperature or total tonnage produced (@fig-biomass-variation).

```{r fig-biomass-variation, class.chunk = "aside"}
plot(coho_biom_summ$temp_c, coho_biom_summ$perc_diff)
# plot(coho_biom_summ$tonnes_per_farm, coho_biom_summ$perc_diff)
```

```{r fig-biomass-produced, class.chunk = "aside"}
#| fig-cap: Modelled vs observed salmon biomass produced

coho_biom %>% 
  ggplot(aes(x = tonnes_per_farm, y = mean, colour = country)) +
  geom_point(size = 2.5) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  scale_x_continuous(limits = c(0,1700)) +
  scale_y_continuous(limits = c(0,1700)) +
  scale_colour_brewer(palette = "Set1") +
  labs(y = "Modelled biomass produced (t)", x = "Observed biomass produced (t)") +
  prettyplot() +
  theme(legend.position = "top", legend.title = element_blank())
```

@@

# Statistical analysis & code availability

All analysis was conducted in R version `r R.version[['major']]`.`r R.version[['minor']]` "`r R.version[['nickname']]`" [@R_base]. 
Major packages used include `r cite_packages(packages)`.
For a full list of R packages used see the lockfile on [Github](https://www.github.com).
This manuscript was written in Quarto [@quarto] using TinyTex [@tinytex] and the acronyms extension [@acronyms].

>>>>

@@

farm_temp_robinson <- st_sf(farm_temp_means) %>% 
  st_transform(crs = "+proj=robin")

p_bigmap_robinson +
  geom_sf(data = farm_temp_robinson, aes(color = mean_temp), size = 1) +
  scale_color_viridis_c(option = "turbo") +
  guides(col = guide_colourbar(
    title = "Mean temperature (°C)",
    direction = "horizontal", position = "bottom",
    label.position = "bottom", title.position = "top", 
    title.vjust = 1, title.hjust = 0.5,
    frame.colour = "black", ticks.colour = "black",ticks.linewidth = 2,
    barwidth = 25, barheight = 1.5
  ))

@@

p_bigmap_temperature <- p_bigmap_mercator +
  geom_sf(data = farm_temp_means, aes(color = mean_temp, geometry = geometry), size = 2) +
  scale_color_viridis_c(option = "turbo") +
  guides(col = guide_colourbar(
    title = "Mean temperature (°C)",
    direction = "vertical", position = "right",
    label.position = "right", title.position = "top", 
    title.vjust = 1, title.hjust = 0.5,
    frame.colour = "black", ticks.colour = "black",ticks.linewidth = 2,
    barwidth = 1.5, barheight = 20
  )) +
  prettyplot() +
  labs(y = "Latitude", x = "Longitude")

p_insets <- get_insets_mercator(p_bigmap_temperature)

grid_1 <- cowplot::plot_grid(
  p_insets[["CAN1"]], p_insets[["CAN2"]], 
  ncol = 2, rel_widths =  c(1, 1.35)
  )
grid_2 <- cowplot::plot_grid(p_Eur,  p_Aus,  nrow = 2, rel_heights = c(1, 1))
grid_3 <- cowplot::plot_grid(grid_2, p_Chi,  ncol = 2, rel_widths =  c(1, 1))

@@

@@

@@
